---
title: "Naïve Bayes/Bayesian Network Exercise"
author: "Etana Disasa"
date: "9/25/2019"
output:
  html_document:
    df_print: paged
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(class)
library(reshape)
library(Rfast)
library(caret)
library(dplyr)
```

# Introduction

#### This exercise has a few questions of solving probabily using Naïve Bayes/Baysian. The Naïve Bayes classifier a probabilty model that is used to do classification based on the Bayes' Theorem. Below are the problems for this assignment. 

# Problems

### PROBLEM 1

#### In a study of pleas and prison sentences, it is reported that 42% of the subjects were sent to prison. Among those set to prison, 38% plead guilty. Among those not sent to prison, 50% plead guilty.

##### Let P = Sent to prison; ¬P = Not sent to prison; and G = Pleaded guilty;
<b>Given: <br>
<i>P(P) = 0.42 <br>
P(P⋂G) = P(P) * 0.38 = 0.42 * 0.38 = `r 0.42*0.38` <br>
P(¬P⋂G) = P(¬P) * 0.50 = 0.58 * 0.50 = `r 0.58*0.5`</i></b> <br>

#### 1. If a subject is randomly selected, what is the probability of getting a person who was not sent to prison. <br>
<b><i>P(¬P) = 1 - P(P) = 1 - 0.42 = 0.58</i></b>

#### 2. If a subject is randomly selected, and it is known that the subject entered a guilty plea, what is the probability that this subject was not sent to prison. <br>
<b>P(¬P|G) = P(¬P⋂G)/P(G) <br>
First, lets caclulate P(G) <br>
<i>P(G) = P(P⋂G) + P(¬P⋂G) <br>
P(G) = `r 0.42*0.38` + `r 0.58*0.5` = `r (0.42*0.38) + (0.58*0.5)` <br>
P(G) = `r (0.42 *0.38)+(0.58*0.5)` </i></b><br><br>
Now, lets calculate P(¬P|G)<br>
<i><b> P(¬P|G) = P(¬P⋂G)/P(G)<br>
P(¬P|G) = `r 0.58*0.5`/`r (0.42*0.38) + (0.58*0.5)` = `r (0.58*0.5)/ ((0.42*0.38) + (0.58*0.5))`</i></b>

#### 3. If a subject is randomly selected, what is the probability of getting someone who was sent to prison.
<b><i>P(P) = 0.42</i></b>

#### 4. If a subject is randomly selected, and it is known that the subject entered a guilty plea, what is the probability that this person was sent to prison.
<b><i>(P|G) = P(P⋂G)/P(G)<br>
(P|G) = `r 0.42*0.38`/`r (0.42*0.38)+(0.58*0.5)`<br>
(P|G) = `r (0.42*0.38)/((0.4*0.38)+(0.58*0.5))`</i></b>

### PROBLEM 2
### Aquiring Data from the given table
```{r}
CustomerID <- c(1:20)
Gender <- c(rep("M",6), rep("F",4), rep("M",4), rep("F",6))
CarType <- c("Family",rep("Sports",8), "Luxury",rep("Family",3), rep("Luxury",7))
ShirtSize <- c("Small",rep("Medium",2), "Large",rep("Extra Large",2), rep("Small",2),"Medium",rep("Large",2),"Extra Large","Medium","Extra Large",rep("Small",2),rep("Medium",3),"Large")
Class <- c(rep("C0",10),rep("C1",10))

MyTable <- as.data.frame(cbind(CustomerID, Gender, CarType, ShirtSize, Class))
MyTable
```


### PROBLEM 3
### <i>A) WHAT IS THE VALUE EACH FOLLOWING PRBABILITY</i>
### Question 1: P(Gender = M | Class = C0)

#### P(M|0) = P(M⋂0)/P(0)
P(M⋂0) = <i>`nrow(filter(MyTable, Gender=="M" & Class=="C0")) / nrow(MyTable)` = `r nrow(filter(MyTable, Gender=="M" & Class=="C0")) / nrow(MyTable)`</i><br>
P(0) = <i>`nrow(filter(MyTable, Class=="C0")) / nrow(MyTable)` = `r nrow(filter(MyTable, Class=="C0")) / nrow(MyTable)`</i>

#### P(M|0) = `r nrow(filter(MyTable, Gender=="M" & Class=="C0")) / nrow(MyTable) / (nrow(filter(MyTable, Class=="C0")) / nrow(MyTable))`


### Question 2: P(Gender = F | Class = C1)

#### P(F|1) = P(F⋂1)/P(1)
P(F⋂0) = <i>`nrow(filter(MyTable, Gender=="F" & Class=="C1")) / nrow(MyTable)` = `r nrow(filter(MyTable, Gender=="F" & Class=="C1")) / nrow(MyTable)`</i><br>
P(1) = <i>`nrow(filter(MyTable, Class=="C1")) / nrow(MyTable)` = `r nrow(filter(MyTable, Class=="C1")) / nrow(MyTable)`</i>

#### P(F|1) = `r nrow(filter(MyTable, Gender=="F" & Class=="C1")) / nrow(MyTable) / (nrow(filter(MyTable, Class=="C1")) / nrow(MyTable))`


### Question 3: P(Car Type = Family | Class = C0) 

#### P(F|0) = P(F⋂0)/P(0)
P(F⋂C) = <i>`nrow(filter(MyTable, CarType=="Family" & Class=="C0")) / nrow(MyTable)` = `r nrow(filter(MyTable, CarType=="Family" & Class=="C0")) /nrow(MyTable)`</i> <br>
P(0)   = <i>`nrow(filter(MyTable, Class=="C0")) / nrow(MyTable)`= `r nrow(filter(MyTable, Class=="C0")) / nrow(MyTable)`</i>

#### P(M|0) = `r nrow(filter(MyTable, CarType=="Family" & Class=="C0")) / nrow(MyTable) / (nrow(filter(MyTable, Class=="C0")) / nrow(MyTable))`


### Question 4: P(Car Type = Family | Class = C1)

#### P(C|1) = P(C⋂1)/P(1)
P(C⋂1) = <i>`nrow(filter(MyTable, CarType=="Family" & Class=="C1")) / nrow(MyTable)` = `r nrow(filter(MyTable, CarType=="Family" & Class=="C1")) /nrow(MyTable)`</i> <br>
P(1)   = <i>`nrow(filter(MyTable, Class=="C1")) / nrow(MyTable)`= `r nrow(filter(MyTable, Class=="C1")) / nrow(MyTable)`</i>

#### P(C|1) = `r nrow(filter(MyTable, CarType=="Family" & Class=="C1")) / nrow(MyTable) / (nrow(filter(MyTable, Class=="C1")) / nrow(MyTable))`


### Question 5: P(Shirt Size = Medium | Class = C0)

#### P(M|0) = P(M⋂0)/P(0)
P(M⋂C) = <i>`nrow(filter(MyTable, ShirtSize=="Medium" & Class=="C0")) / nrow(MyTable)` = `r nrow(filter(MyTable, ShirtSize=="Medium" & Class=="C0")) /nrow(MyTable)`</i> <br>
P(0)   = <i>`nrow(filter(MyTable, Class=="C0")) / nrow(MyTable)`= `r nrow(filter(MyTable, Class=="C0")) / nrow(MyTable)`</i>

#### P(M|0) = `r nrow(filter(MyTable, ShirtSize=="Medium" & Class=="C0")) / nrow(MyTable) / (nrow(filter(MyTable, Class=="C0")) / nrow(MyTable))`


### Question 6: P(Shirt Size = Medium | Class = C1)

#### P(S|1) = P(S⋂1)/P(1)
P(S⋂1) = <i>`nrow(filter(MyTable, ShirtSize=="Medium" & Class=="C1")) / nrow(MyTable)` = `r nrow(filter(MyTable, ShirtSize=="Medium" & Class=="C0")) /nrow(MyTable)`</i> <br>
P(1)   = <i>`nrow(filter(MyTable, Class=="C1")) / nrow(MyTable)`= `r nrow(filter(MyTable, Class=="C1")) / nrow(MyTable)`</i>

#### P(S|1) = `r nrow(filter(MyTable, ShirtSize=="Medium" & Class=="C1")) / nrow(MyTable) / (nrow(filter(MyTable, Class=="C1")) / nrow(MyTable))`

### <I>B) USE NAÏVE BAYE'S CLASSIFIER TO FIND THE CLASS OF: </i>
### P(Gender = F, Car Type = Family, Shirt Size = Medium)

P(Gender = F, Car Type = Family, Shirt Size = Medium) = P(Gender = F)*P(Car Type = Family)*P(Shirt Size = Medium)<br>
P(Gender = F, Car Type = Family, Shirt Size = Medium) = 10/20 * 1/20 * 8/20 = 0.01

# Conclusion
#### For this assignment, we have used the Bayes' theorem to calculate the probabilities for the data presented. In the following excersise, we will apply this theorem in order to do a classification. 

# References
#### https://sebastianraschka.com/Articles/2014_naive_bayes_1.html 
#### https://rpubs.com/riazakhan94/naive_bayes_classifier_e1071
