---
title: "Decision Tree Exercise"
author: "Etana Disasa"
date: "09/28/2019"
output:
  html_document:
    df_print: paged
  word_document: default
---

```{r echo=FALSE, warning = FALSE, message = FALSE}
library(rpart)
library(rattle)
library(rpart.plot)
library(caret)
library(partykit)
library(randomForest)
```

# Introduction
##### In today's exercise, we will use Decision Tree and Random Forest classifiers in order to determine the quality of wine. We will utilize a dataset from the UCI.edu website which I have downloaded and attached along with this assignment. The purpose of this assignment is to see which classifier classifies the quality of wine more accurately than the other. 

# Method
### Acquiring and Exploring Data
##### In order to make sure we have a working dataset when there is no internet connection, we would need to download the original dataset to our machine. I have also commented (#ed) the code line with the original dataset url. 
```{r}
# MyData <- read.csv(file="https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv", header = TRUE, sep = ";")
MyWine <- read.csv(file = "winequality-red.csv", header = TRUE, sep = ";")
str(MyWine)
```

##### This dataset contains 1599 observations with 12 variables. There are six different wine qualities. All 12 variables are numerical (numbers and integers). We would need to change the predictor variable into a factor. And also, let us explore how many of the wines in our dataset belong to a particular wine quality. 

##### We may also give an arbitrary quality for the wines such as, 3 and 4 as `Dad`; 5 and 6 as `OK`, 7 and 8 as `Good`. That was we can offer a wine quality in these three descriptors. 

```{r}
MyWine$quality[MyWine$quality == "3"] <- "Bad"
MyWine$quality[MyWine$quality == "4"] <- "Bad"
MyWine$quality[MyWine$quality == "5"] <- "OK"
MyWine$quality[MyWine$quality == "6"] <- "OK"
MyWine$quality[MyWine$quality == "7"] <- "Best"
MyWine$quality[MyWine$quality == "8"] <- "Best"
str(MyWine)
MyWine$quality <- factor(MyWine$quality)

table(MyWine$quality)
```

### Splitting Data into Train and Test Sets

##### We will use the `rpart()` function to randomly sample 8:2 for training and testing sets. 

```{r}
MySample <- sample(1599, round(1599*0.8, digits = 0))
str(MySample)
MyTrain <- MyWine[MySample, ]
MyTest <- MyWine[-MySample, ]
nrow(MyTrain)
nrow(MyTest)
```

##### Now we have `MyTrain` and `MyTest` datasets, with `r nrow(MyTrain)` and r nrow(MyTest)` observations consecutively according to 8:2 ratio we would like to attain. 

### Decision Tree
##### Below, I have used three different kinds of visualization to help us understand how our decision tree is working. 
```{r}
tm <- rpart(quality~., MyTrain, method = "class")
rpart.plot(tm, tweak = 1.6)
prp(tm)
fancyRpartPlot(tm)
```

### Prediction
```{r}
pred <- predict(tm, MyTest, type = "class")
table(MyTest$quality, pred)
```
##### Using our model, we now no prediction in order to classify the quality of wine. 
### Confusion Matrix
```{r}
confusionMatrix(table(pred, MyTest$quality))
```
##### Our decision tree predicts with approximately 85% accuracy rate. Now let us see how a randomeForest does in predicting win quality.

### Random Forest
##### Now using the same train and test data, lets use Random Forest classification. 

```{r}
tm <- randomForest(quality~., data = MyTrain, importance = TRUE)
```
##### Prediction
```{r}
pred <- predict(tm, MyTest, type = "class")
table(MyTest$quality, pred)
```
### Confusion Matrix
```{r}
confusionMatrix(table(pred, MyTest$quality))
```
##### Here, our random forest predicts with an approximately 88% of accuracy which made our model better. 


# Conclusion
##### Decision trees and Random Forests are classifiers that we had used for this assignment. Given this dataset, random forest predicted with more accuracy than a decision tree model. Given these curcumstances, it would be plausible to suggest to use random forest model in order to predict the quality of wines if the dataset comes with the same variables. 

# References
https://www.r-bloggers.com/predicting-wine-quality-using-random-forests/ <br>
https://rstudio-pubs-static.s3.amazonaws.com/318409_f3e4e00fb69f4e22b82c0e0f1f6014df.html <br>
